---
title: "Assignment5.2"
author: "NgWM"
date: "2025-12-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Commit containing full submission

1. Code for reading in the dataset and/or processing the data

```{r}
old.dir <- getwd()
setwd("C:/Users/User/OneDrive/Desktop Nvidia/DS notes/Chapter 5/Work 5.2")

data <- read.csv("activity.csv")
str(data)
data$date <- as.Date(as.character(data$date))
data$interval <- as.integer(data$interval)
str(data)
summary(data) # to check if there is NA
```

2. Histogram of the total number of steps taken each day
```{r}
activity_noNA <- subset(data,!is.na(steps))
steps_per_day <- aggregate(steps~date, data=activity_noNA, sum)
#plot
hist(steps_per_day$steps,
     main= "Total number of steps per day",
     xlab= "Steps per day",
     col = "lightblue")

```

3. Mean and median number of steps taken each day
```{r}
mean(steps_per_day$steps)
median(steps_per_day$steps)
```
4. Time series plot of the average number of steps taken
```{r}
steps_per_interval <- aggregate(steps~interval, data=activity_noNA,mean)
#plot
plot(steps_per_interval$interval, steps_per_interval$steps,
     type="l",
     xlab= "5-min interval",
     ylab= "Average number of steps taken",
     main=" Average daily Step Activity")
```
5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max_interval_value <- which.max(steps_per_interval$steps)
steps_per_interval$interval[max_interval_value]
```
6. Code to describe and show a strategy for imputing missing data
```{r}
total.NA <-  sum(is.na(data$steps))
interval_means <- aggregate(steps~interval,data=data,FUN=mean,na.rm=TRUE)
#Merge
activity_imputed <- merge(data,interval_means, 
                          by = "interval",
                          suffixes = c("",".mean"))
# If steps is NA, repalce by interval mean
activity_imputed$steps[is.na(activity_imputed$steps)] <- activity_imputed$steps.mean[is.na(activity_imputed$steps)]
# Drop hoelper column
activity_imputed_steps.mean <- NULL
str(activity_imputed)
sum(is.na(activity_imputed$steps))
```
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
steps_per_day_imputed <-  aggregate (steps~date, data=activity_imputed, sum)

hist(steps_per_day_imputed$steps,
     main = " Total steps per day (after imputation)", 
     xlab = " Steps per day",
     col  = "lightgreen")
```
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
# create factor variable indicating weekdays & weekends
day_type <- function(d){
 wd <- weekdays(d)
 if (wd %in% c("Saturday","Sunday")) "weekend" else "weekday"
}  

activity_imputed$day_type <- factor(sapply(activity_imputed$date,day_type),levels = c("weekday","weekend"))

# average steps per interval for weekday vs weekends
steps_interval_daytype <- aggregate(steps~interval+day_type, 
                                    data= activity_imputed, FUN = mean)

par(mfrow = c(2,1), mar =c(4,4,2,1))

# plot
with(subset(steps_interval_daytype,day_type =="weekday"), 
     plot(interval, steps, 
          type = "l", 
          main ="Weekdays",
          xlab = "5-minute interval",
          ylab =" Average steps"))

with(subset(steps_interval_daytype,day_type =="weekend"), 
     plot(interval, steps, 
          type = "l", 
          main ="Weekends",
          xlab = "5-minute interval",
          ylab =" Average steps"))
```

This is the assigment for Chapter 5.2 Coursera From JHU. 

